---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Travis-CI Build Status](https://travis-ci.org/robinsones/funneljoin.svg?branch=master)](https://travis-ci.org/robinsones/funneljoin)

The goal of funneljoin is to make it easy to analyze behavior funnels. For example, maybe you're interested in finding the people who visit a page and then register. Or you want all the times people click on an item and add it to their cart within 2 days. These can all be answered quickly with funneljoin's `after_join()` or `funnel_start()` and `funnel_step()`. As funneljoin uses dplyr, it can also work with remote tables, **but has only been tried on postgres**. 

For more examples of how to use funneljoin, check out [the vignette](https://robinsones.github.io/funneljoin/articles/funneljoin.html), which shows different types of joins and the optional arguments, or this [blog post](https://hookedondata.org/introducing-the-funneljoin-package/), which showcases how to use funneljoin analyze questions and answers on StackOverflow. 

Funneljoin was developed at [DataCamp](https://www.datacamp.com/) by Anthony Baker, David Robinson, and Emily Robinson and continues to be maintained primarily by Emily.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("robinsones/funneljoin")
```

## after_join()

```{r, message = FALSE}
library(dplyr)
library(funneljoin)
```

We'll take a look at two tables that come with the package, `landed` and `registered`. Each has a column `user_id` and `timestamp`.

Let's say we wanted to get the first time people landed and the first time afterward they registered. We would `after_inner_join()` with a `first-firstafter` type: 

```{r}
landed %>%
  after_inner_join(registered, 
                   by_user = "user_id",
                   by_time = "timestamp",
                   type = "first-firstafter",
                   suffix = c("_landed", "_registered"))
```

The first two arguments are the tables we're joining, with the first table being the events that happen first. We then specify: 

  - `by_time`: the time columns in each table. This would typically be a datetime or a date column. These columns are used to filter for time y being after or the same as time x.
  - `by_user`:the user or identity columns in each table. These must be identical for a pair of rows to match.
  - `type`: the type of funnel used to distinguish between event pairs, such as "first-first", "last-first", "any-firstafter".
  - `suffix` (optional): just like dplyr’s join functions, this specifies what should be appended to the names of columns that are in both tables.

`type` can be any combination of `first`, `last`, `any`, and `lastbefore` with `first`, `last`,  `any`, and `firstafter`. Some common ones you may use include:  

  - **first-first**: Take the earliest x and y for each user **before** joining. For example, you want the first time someone entered an experiment, followed by the first time someone **ever** registered. If they registered, entered the experiment, and registered again, you do not want to include that person. 
  - **first-firstafter**: Take the first x, then the first y after that. For example, you want when someone first entered an experiment and the first course they started afterwards. You don't care if they started courses before entering the experiment. 
  - **lastbefore-firstafter**: First x that's followed by a y before the next x. For example, in last click paid ad attribution, you want the last ad someone clicked before the first subscription they did afterward. 
  - **any-firstafter**: Take all Xs followed by the first Y after it. For example, you want all the times someone visited a homepage and their first product page they visited afterwards.
  - **any-any**: Take all Xs followed by all Ys. For example, you want all the times someone visited a homepage and **all** the product pages they saw afterward. 

If your time and user columns have different names, you can work with that too: 

```{r}
landed <- landed %>%
  rename(landed_at = timestamp,
         user_id_x = user_id)

registered <- registered %>%
  rename(registered_at = timestamp,
         user_id_y = user_id)
```

```{r}
landed %>%
  after_inner_join(registered, 
                   by_user = c("user_id_x" = "user_id_y"),
                   by_time = c("landed_at" = "registered_at"),
                   type = "first-first")
```

## funnel_start() and funnel_step()

Sometimes you have all the data you need in one table. For example, let's look at this table of user activity on a website. 

```{r}
activity <- tibble::tribble(
  ~ "user_id", ~ "event", ~ "timestamp",
  1, "landing", "2019-07-01",
  1, "registration", "2019-07-02",
  1, "purchase", "2019-07-07",
  1, "purchase", "2019-07-10",
  2, "landing", "2019-08-01",
  2, "registration", "2019-08-15",
  3, "landing", "2019-05-01",
  3, "registration", "2019-06-01",
  3, "purchase", "2019-06-04",
  4, "landing", "2019-06-13"
)
```

We can use `funnel_start()` and `funnel_step()` to make an activity funnel. `funnel_start()` takes five arguments: 

  - `tbl`: The table of events.
  - `moment_type`: The first moment, or event, in the funnel.
  - `moment`: The name of the column that indicates the `moment_type`.
  - `tstamp`: The name of the column with the timestamps of the moment. 
  - `user`: The name of the column indicating the user who did the moment.

```{r}
activity %>%
  funnel_start(moment_type = "landing", 
               moment = "event", 
               tstamp = "timestamp", 
               user = "user_id")
```

`funnel_start()` returns a table with the user_ids and a column with the name of your timestamp column, `_`, and the moment type. This table also includes metadata.  

To add more moments to the funnel, you use `funnel_step()`. Since you've indicated in `funnel_start()` what columns to use for each part, now you only need to have the `moment_type` and the `type` of `after_join()` (e.g. "first-first", "first-any"). 

```{r}
activity %>%
  funnel_start(moment_type = "landing", 
               moment = "event", 
               tstamp = "timestamp", 
               user = "user_id") %>%
  funnel_step(moment_type = "registration",
              type = "first-firstafter")
```

You can continue stacking on `funnel_step()` with more moments. 

```{r}
activity %>%
  funnel_start(moment_type = "landing", 
               moment = "event", 
               tstamp = "timestamp", 
               user = "user_id") %>%
  funnel_step(moment_type = "registration",
              type = "first-firstafter") %>%
  funnel_step(moment_type = "purchase",
              type = "first-any")
```

If you use a `type` that allows multiple moments of one type for a user, like "first-any", you will get more rows per user rather than more columns. For example, user 1 had two purchases, so she now has two rows. The `timestamp_landing` and `timestamp_registration` is the same for both rows, but they have a different `timestamp_purchase`. 

Finally, you can use the `summarize_funnel()` to understand how many and what percentage of people make it through to each next step of the funnel. We can also switch to `funnel_steps()` to shorten our code a bit - we give it a character vector of `moment_types` in order and the `type` for each step. 

```{r}
activity %>%
  funnel_start(moment_type = "landing", 
               moment = "event", 
               tstamp = "timestamp", 
               user = "user_id") %>%
  funnel_steps(moment_types = c("registration", "purchase"),
              type = "first-firstafter") %>%
  summarize_funnel()
```

`nb_step` is how many users made it to each step, `pct_cumulative` is what percent that is out of the original step, and `pct_step` is what percentage that is out of those who made it to the previous step. So in our case, 2 people had a purchase, which is 50% of the people who landed but 66% of those who registered.

## Reporting bugs and adding features

If you find any bugs or have a feature request or question, please [create an issue](https://github.com/robinsones/funneljoin/issues/new). If you'd like to add a feature, tests, or other functionality, please also make an issue first and let's discuss! 
