---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Travis-CI Build Status](https://travis-ci.org/robinsones/funneljoin.svg?branch=master)](https://travis-ci.org/robinsones/funneljoin)

The goal of funneljoin is to make it easy to analyze behavior funnels. For example, maybe you're interested in finding the people who visit a page and then register. Or you want all the times people click on an item and add it to their cart within 2 days. 
You can do this with funneljoin's `after_join()` function. The arguments are: 

  - `x`: a dataset with the first set of behaviors.
  - `y`: a dataset with the second set of behaviors.
  - `by_time`: a character vector to specify the time columns in x and y. This would typically be a datetime or a date column. These columns are used to filter for time y being after time x.
  - `by_user`: a character vector to specify the user or identity columns in x and y.
  - `mode`: the method used to join: "inner", "full", "anti", "semi", "right", "left". You can also use `after_mode_join` instead (e.g. `after_inner_join` instead of `after_join(..., mode = "inner")`)
  - `type`: the type of funnel used to distinguish between event pairs, such as "first-first", "last-first", "any-firstafter". See types of funnels for details.
  - `max_gap` (optional): the maximum gap between events. Can be a integer representing the number of seconds between events or a difftime object. 
  - `gap_col` (optional): whether to return a numeric column, `.gap`, with the time difference in seconds between the events. Defaults to `FALSE`. 

As funneljoin uses dplyr, it can also work with remote tables, **but has only been tried on postgres**. Some of the functionality, especially the gap joins, use SQL code that may not work with different versions. 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("robinsones/funneljoin")
```

## Types of funnels

`type` can be any combination of `first`, `last`, `any`, and `lastbefore` with `first`, `last`,  `any`, and `firstafter`. Some common ones you may use include:  

  - **first-first**: Take the earliest x and y for each user **before** joining. For example, you want the first time someone entered an experiment, followed by the first time someone **ever** registered. If they registered, entered the experiment, and registered again, you do not want to include that person. 
  - **first-firstafter**: Take the first x, then the first y after that. For example, you want when someone first entered an experiment and the first course they started afterwards. You don't care if they started courses before entering the experiment. 
  - **lastbefore-firstafter**: First x that's followed by a y before the next x. For example, in last click paid ad attribution, you want the last ad someone clicked before the first subscription they did afterward. 
  - **any-firstafter**: Take all Xs followed by the first Y after it. For example, you want all the times someone visited a homepage and their first product page they visited afterwards.
  - **any-any**: Take all Xs followed by all Ys. For example, you want all the times someone visited a homepage and **all** the product pages they saw afterward. 

See the vignette for more examples. 

### Example

```{r}
library(dplyr)
library(funneljoin)
```

```{r}
landed <- tibble::tribble(
  ~user_id, ~timestamp,
  1, "2018-07-01",
  2, "2018-07-01",
  3, "2018-07-02",
  4, "2018-07-01",
  4, "2018-07-04",
  5, "2018-07-10",
  5, "2018-07-12",
  6, "2018-07-07",
  6, "2018-07-08"
) %>%
  mutate(timestamp = as.Date(timestamp))

registered <- tibble::tribble(
  ~user_id, ~timestamp,
  1, "2018-07-02",
  3, "2018-07-02",
  4, "2018-06-10",
  4, "2018-07-02",
  5, "2018-07-11",
  6, "2018-07-10",
  6, "2018-07-11",
  7, "2018-07-07"
) %>%
  mutate(timestamp = as.Date(timestamp))
```

Let's say we wanted to get only the first time people landed and registered. We would use a `first-first inner` join: 

```{r}
landed %>%
  after_inner_join(registered, 
                   by_user = "user_id",
                   by_time = "timestamp",
                   type = "first-first")
```

If your time and user columns have different names, you can work with that too: 

```{r}
landed <- landed %>%
  rename(landed_at = timestamp,
         user_id_x = user_id)

registered <- registered %>%
  rename(registered_at = timestamp,
         user_id_y = user_id)
```

```{r}
landed %>%
  after_inner_join(registered, 
                   by_user = c("user_id_x" = "user_id_y"),
                   by_time = c("landed_at" = "registered_at"),
                   type = "first-first")
```

## Summarizing funnels

Funneljoin also contains two functions to summarize funnels: `summarize_conversions()` and `summarize_prop_tests()`- see the vignette for details. 

## Reporting bugs and adding features

If you find any bugs or have a feature request or question, please [create an issue](https://github.com/datacamp/funneljoin/issues/new). If you'd like to add a feature, tests, or other functionality, please also make an issue first and let's discuss! 
